ZPY: Open Source Synthetic Data for Computer Vision

### 2021-07-15 docs/paper/paper.tex
mvp2 updates

\begin{abstract}
Synthetic data presents a unique solution to the huge data requirements of computer vision with deep learning. In this work, we present zpy \footnote{All code is available on GitHub at \url{http://github.com/ZumoLabs/zpy}}, an open source framework for creating synthetic data. Built on top of the popular open source 3D platform Blender, zpy is designed with accessibility and readability in mind. Open source synthetic data toolkits like zpy are the bridge between the more mature tools of the 3D workflow and the machine learning frameworks. We make the case for why open source synthetic data is important to solve issues such as fairness and bias by democratizing access to data.
\end{abstract}

@@

The fourth perceivable dimension of our reality is time. Asset scripting is the process of defining the behaviors of assets within scenes over time. One type of asset scripting is called animation, which consists of creating sequential mesh deformations that create the illusion of natural movement. Animation is a tedious manual task because an artist must define every frame; expert animators spend decades honing their digital puppeteering skills. Specialized software is often used to automate this task as much as possible, and technologies such as Motion Capture (MoCap) can be used to record the movement of real objects and play those movements back on virtual assets.

Game Engines are software tools that allow for more structured and systematic asset scripting, mostly by providing software interfaces (e.g., code) to control the virtual world. Used extensively in the video game industry after which they were named, examples include Unity \cite{unity3d}, Unreal Engine \cite{unrealengine}, GoDot \cite{godot}, and Roblox \cite{roblox}. These game engines support rule-based spawning, animation, and complex interactions between assets in the virtual world. Programming within game engines is a separate skillset to modeling and animating and is usually done by separate engineers within an organization.

@@

\subsection{Democratization of Data}

Modern computer vision systems use learning based methods and thus require large and diverse datasets. These datasets are almost exclusively collected: images are stored in databases as they are cumulatively generated by users of a product over time. Annotations are created by hand, usually by a third party provider using low-skilled labor in third world countries. Collecting and labeling a dataset large enough to train a robust computer vision model can take years. Only companies that have the scale and have set up the infrastructure to collect and store large datasets will have the datasets required to train models.

In order to compete, small or newly formed companies often resort to purchasing data from a third party supplier. This market for the selling and reselling of collected data presents an existential threat to privacy. Though some regulations have emerged, famously GDPR in Europe, these have yet to change the landscape of the market for collected data. Synthetic data can democratize access to large-scale datasets by reducing the time required to collect these datasets and eliminating the cost of labeling these datasets.

@@

The large cost to collect and annotate datasets makes it prohibitive for most small organizations to create their own datasets, and as such it is common practice to use one of a small selection of openly available datasets (such as ImageNet). However, because of how these datasets have been collected over time, and due to the unequal global distribution of technologies such as cell phone cameras, these datasets are often biased. The real world is biased and unfair, and these collected datasets do not represent the large variety of cultural, demographic, or gender diversity  in human datasets \cite{shankar2017classification}. These datasets do not capture all geographic differences in object recognition datasets \cite{DBLP:journals/corr/abs-1906-02659}. The nature of statistical learning methods means that training on these biased datasets will result in a biased model: a model is only as good as the data on which it is trained. As such, it follows that the best way to reduce bias in modern computer vision systems is to improve the datasets these systems are trained on.

Synthetic datasets offer full control of the distribution and can thus be designed to be more representative. A synthetic human dataset can equally represent cultural, demographic, or gender diversity. Synthetic datasets as a method of reducing bias has been explored in a variety of computer vision domains \cite{DBLP:journals/corr/abs-2004-13866}.

@@

Abstracting away the difficulties of the cloud workflow and providing an intuitive and convenient interface for parallelizing computation is thus important for the synthetic data workflow. Dataset generation jobs can be made with the zpy Python API \ref{lst:api} and CLI \ref{lst:cli} such that many computers are used, thus speeding up the generation job. The ZumoLabs cloud uses orchestration technology, e.g. Kubernetes, to spin up and manage new image generation nodes.

@@

The zpy python package is organized as a collection of seperate modules. These modules are separated based on their dependencies and functionality. This makes it easier for end users to add custom functionality, or ignore modules with dependencies they do not want. This type of modular design can be seen in other popular python libraries, such as Matplotlib \cite{Hunter:2007}.

>>>>

\begin{abstract}
Synthetic data presents a unique solution to the huge data requirements of computer vision with deep learning. In this work, we present zpy \footnote{All code is available on GitHub at \url{http://github.com/ZumoLabs/zpy}}, an open source framework for creating synthetic data. Built on top of the popular open source 3D toolset Blender, zpy is designed with accessibility and readability in mind. Open source synthetic data toolkits like zpy are the bridge between the more mature tools of the 3D workflow and the machine learning frameworks. We make the case for why open source synthetic data is important to solve issues such as fairness and bias by democratizing access to data.
\end{abstract}

@@

The fourth perceivable dimension of our reality is time. Asset scripting is the process of defining the behaviors of assets within scenes over time. One type of asset scripting is called animation, which consists of creating sequential mesh deformations that create the illusion of natural movement. Animation is a tedious manual task because an artist must define almost every frame; expert animators spend decades honing their digital puppeteering skills. Specialized software is often used to automate this task as much as possible, and technologies such as Motion Capture (MoCap) can be used to record the movement of real objects and play those movements back on virtual assets.

Game Engines are software tools that allow for more structured and systematic asset scripting, mostly by providing software interfaces (e.g., code) to control the virtual world. Used extensively in the video game industry after which they were named, examples include Unity \cite{unity3d}, Unreal Engine \cite{unrealengine}, Godot \cite{godot}, and Roblox \cite{roblox}. These game engines support rule-based spawning, animation, and complex interactions between assets in the virtual world. Programming within game engines is a separate skillset to modeling and animating and is usually done by separate engineers within an organization.

@@

\subsection{Democratization of Data}
\label{sec:democratizationofdata}

Modern computer vision systems use learning based methods and thus require large and diverse datasets. The need for large and diverse datasets results in a data moat that precludes small companies from competing in the CV market. To amass a good dataset, a company needs time and money: precisely the two things that small nascent companies do not have.

\subsection{The Status Quo}

The problem lies in how the data pipeline works today. Currently, datasets are almost exclusively collected. mages are stored in databases as they are cumulatively generated by users of a product over time. Annotations are created by hand, usually by a third party provider using low-skilled labor in third world countries. Collecting and labeling a dataset large enough to train a robust computer vision model can take years. Moreover, labeling is not a one-off step, but rather a process that has to be repeated many times. Hand labeled data has shown itself to be imperfect. As one example, studies have found that the widely used ImageNet test set has an estimated label error rate of 5.8\% \cite{dataerrors}. The result is that only companies that have set up the infrastructure to collect and store large datasets, and then the money to subsequently label them, will have the datasets required to train models., Often, the only recourse for a small or newly formed company is to  purchase data from a third party supplier.

@@

The large cost to collect and annotate datasets makes it prohibitive for most small organizations to create their own datasets, and as such it is common practice to use one of a small selection of openly available datasets (such as ImageNet). This practice has resulted in the pervasive phenomenon of biased algorithms that perpetuate and codify racism, sexism, and bigotry. In the case of human datasets, irresponsibly collected datasets can also raise serious privacy concerns.

\subsection{Algorithmic Bias}

Algorithmic bias is when an algorithm exhibits a prejudice or inclination for one outcome over another. Because of how openly available datasets have been collected over time, and due to the unequal global distribution of technologies such as cell phone cameras, datasets are often biased.

Algorithms don’t make fairness; they automate the status quo. The real world is biased and unfair. In human datasets, inevitably, publicly available image datasets exhibit Amero-centric and Euro-centric representation bias and do not adequately address true cultural, demographic, or gender diversity\cite{shankar2017classification}. Widely used public datasets for object recognition also fail to capture all geographic differences\cite{DBLP:journals/corr/abs-1906-02659}.

Biased datasets result in biased algorithms. The nature of statistical learning methods means that training on these biased datasets will result in a biased model since a model is only as good as the data on which it is trained. The pervasive use of data-driven algorithms has complex social implications: from racial facial recognition (citation: https://nvlpubs.nist.gov/nistpubs/ir/2019/NIST.IR.8280.pdf), to predictive policing systems that act to confirm and deepen existing biases (citation: https://www.technologyreview.com/2020/07/17/1005396/predictive-policing-algorithms-racist-dismantled-machine-learning-bias-criminal-justice/). 
\subsection{Privacy}

Using real data to train models presents very serious privacy risks with the rise of comprehensive privacy regimes like the GDPR in Europe and the CCPA in California. Openly available datasets offer no privacy guarantees as to whether all humans and biometrics represented in the dataset have been acquired with consent.

Generally, today, the practice when using real datasets is to subject them to a de-identification process. De-identification, sometimes referred to as anonymization, strips a dataset of personal identifiers. However, there are two problems with this process: one, all anonymized data is subject to reversal. The only real bar is the state of technology at the point in time. Anonymized data today becomes pseudonymized data (data that can be reversed and re-identify individuals) tomorrow as models become better at re-identifying data points. Second, anonymized data is not as useful: it is less rich and many times lacks the information required to train a model.

\subsection{Synthetic Data as a Solution}

Synthetic data can democratize access to large-scale datasets by reducing the time required to collect these datasets and eliminating the cost of labeling these datasets.

Furthermore, synthetic data presents a solution to algorithmic bias: synthetic datasets offer complete control of the distribution and can thus be representatively designed. A synthetic human dataset can be constructed to represent cultural, demographic, and gender diversity equally.  A variety of computer vision domains have explored using synthetic datasets to reduce bias effectively \cite{DBLP:journals/corr/abs-2004-13866}.

Finally, synthetic datasets, by virtue of being created and not collected, guarantees privacy compliance. Synthetic data is free of personal identifiers that could be susceptible to re-identification down the road. Synthetic data guarantees privacy by changing the paradigm and getting rid of any need to use real data.

@@

Abstracting away the difficulties of the cloud workflow and providing an intuitive and convenient interface for parallelizing computation is thus important for the synthetic data workflow. Dataset generation jobs can be made with the zpy Python API \ref{lst:api} and CLI \ref{lst:cli} such that many computers are used, thus speeding up the generation job. The Zumo Labs cloud uses orchestration technology, e.g. Kubernetes, to spin up and manage new image generation nodes.

@@

The zpy python package is organized as a collection of separate modules. These modules are separated based on their dependencies and functionality. This makes it easier for end users to add custom functionality, or ignore modules with dependencies they do not want. This type of modular design can be seen in other popular python libraries, such as Matplotlib \cite{Hunter:2007}.
